---
title: "Median Income"
author: "Umaira"
execute: 
  echo: false
  message: false
  warning: false
---
Get Some more data

```{r}
#| echo: false
#| message: false
#| warning: false

options(tigris_use_cache = TRUE)
options(tigris_progress = FALSE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(tidycensus))
suppressPackageStartupMessages(library(sf))

# Download median household income by census tract for Massachusetts
ma_income <- get_acs(
  geography = "tract",
  variables = "B19013_001",  # Median household income variable
  state = "MA",
  year = 2022,  # Most recent ACS 5-year estimates
  geometry = TRUE  # Set to TRUE if you want spatial data
)
```

```{r}
# Create choropleth map
ggplot(data = ma_income, aes(fill = estimate)) +
  geom_sf(color = "white", size = 0.1) +
  scale_fill_viridis_c(
    option = "magma", 
    labels = scales::dollar_format(),
    name = "Median Income"
  ) +
  theme_minimal() +
  labs(
    title = "Median Household Income by Census Tract in Massachusetts",
    caption = "Source: ACS 5-Year Estimates (2022)"
  ) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )

```

The code above retrieves and maps median household income data for Massachusetts census tracts using the tidycensus and sf packages in R. It pulls the latest available data from the 2022 American Community Survey (5-Year Estimates) through the Census API. By setting geometry = TRUE, it also fetches the geographic boundaries for each tract, enabling spatial visualization. The resulting data is then used to create a choropleth map using ggplot2, where each tract is shaded based on its median income. This map helps reveal income variation across different parts of the state and provides insights into geographic patterns of wealth and inequality.

